% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{llncsdoc}
\usepackage{algorithmic}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{cite}
\usepackage{amsmath}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{hyperref}
\usepackage[parfill]{parskip} 
\usepackage{float}
\usepackage{caption}
%\restylefloat{table}

	%\restylefloat{table}
%
\begin{document}
\pagestyle{empty}
%\centerline{\bf UNGIVERSITY OF BUCHAREST}

%\centerline{\bf FACULTY OF MATHEMATICS AND COMPUTER SCIENCE}

%
%
\title{Data Structures and Algorithms}
%
\titlerunning{Data Structures and Algorithms}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Drago\c{s} Alin Rotaru}
%TODO - nume final
\authorrunning{and D.A.Rotaru} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{}
%
\institute{Computer Science, University of Bucharest, Romania\\
}


\maketitle              % typeset the title of the contribution


\newpage
%----------------------------------------------------------------
%----------------------------------------------------------------
%----------------------------------------------------------------
%----------------------------------------------------------------
%----------------------------------------------------------------

\section{Introduction}
\label{sec:intro}

These seminar notes contain my overview of the Data Structures and Algorithms course held at University of Bucharest. Because the course is based on heavy theoretic lectures, I tried a more practical approach to present some of the notions by discussing problems which arise natural from the main course.

Most of the problems come from a romanian website specialized on programming contests as well as codeforces or topcoder \cite{website:infoarena, website:topcoder, website:codeforces}. Of course, there are more interesting problems to tackle, but unfortunately I limit to the course material although sometimes I will talk about some ad-hoc problems.


\section{Seminar I}
\label{sec:seminar1}

Synthesise first 2 courses:
\begin{itemize}
	\item Basic notions of time and memory complexity.
	\item Stacks and Queues.
\end{itemize}

\subsection{Sketch}
	What is an algorithm? How can we measure the time complexity of a program? Examples (Choosing every pair of elements and erathosthene sieve).
	Introduction to stacks and queues. Details about their implementation and a short tutorial in STL. Can also talk about circular queues and double ended queues.

\subsection{Checking if an expression has brackets in right order}

\subsection{Emulate a queue using 2 stacks}

\subsection{Editor \cite{website:infoarena/editor}}

\subsection{Alee \cite{website:infoarena/alee}}

\subsection{Trompeta \cite{website:infoarena/trompeta}}
Given N digits, find the maximum number with K digits such that the digits preserve the initial order.

\subsection{Tsunami \cite{website:infoarena/tsunami}}
Given a matrix NxM with digits less than a value $K (<= 10)$. Find the number of cells with value less than a threshold $T (T <= K)$ such that when you start from 0 you always go up to digits $<= T$.

\subsection{Add or Multiply}

Take a number X. You can add, subtract, multiply it by numbers in a set $S$. Find the shortest number of steps required to reach from $X$ to $Y$.

\subsection{Take-Out \cite{website:edu-pl:take-out}}

N blocks, each black or white. You can remove at a time $k$ white blocks and $1$ black block such that there is no gap between the removed blocks.
\subsection{Devices}
	You are given a row of n devices, each consuming some subset of k<=8 different resources when turned on, and producing some amount of energy when turned on. For each l from 1 to n you need to find the smallest r such that it's possible to turn on some devices from the segment [l;r] such that no two devices turned on consume the same resource, and that the total energy of the devices turned on is at least z \cite{website:petr1}.

\section{Seminar II}
\label{sec:seminar2}

\begin{itemize}
	\item Divide and conquer, merge-sort, estimating complexity
\end{itemize}

\subsection{Inv \cite{website:infoarena/inv}}
\subsection{Stergeri \cite{website:infoarena/stergeri}}
\subsection{Muzeu \cite{website:infoarena/muzeu}}
\subsection{Binar \cite{website:infoarena/binar}}


\section{Seminar III}

Binary search. Fast exponentiation. K'th element. Time complexity proofs (at least sketches).

\subsection{Classic Task \cite{website:infoarena/classictask}}
\subsection{Nrtri \cite{website:infoarena/nrtri}}
\subsection{Sdo \cite{website:infoarena/sdo}}
	As a further challenge, consider that you can't store $O(N)$ elements. Can you do dynamic medians?
\subsection{Loto \cite{website:infoarena/loto}}

\section{Seminar IV}

Heaps. Binary Heaps. Fibonacci Heaps.

Basic heap operations (Insert, Extract-Min, Merge).

Time, space complexity analysis.
Heap reconstruction in $O(N)$.

\subsection{Sdo \cite{website:infoarena/sdo}}
Can you find the median using heaps? Can you do this dinamically?

\subsection{Merge $K$ lists \cite{website:leetcode/merge}}
You have $K$ sorted arrays. Can you merge them quickly?

\subsection{Sort using heaps}
Classic sort algorithm, this time using heaps. Almost sorted list, can you sort it more efficiently than classic?


\section{Seminar V}

Stanford exercises. \cite{website:stanford/binarytrees}

Wiki page about AVL. \cite{website:wiki:avl}

Get K'th element from a BST. Construct BST from an array.


\section{Seminar VI}

\subsection{Huffman Coding}

\textbf{General problem}:

Given an alphabet $\mathcal{A} = \{a_1, \dots, a_n\}$ and a set of weights $\mathcal{W} = \{w_1, \dots, w_n \}$ compute a set of codewords $\mathcal{C} = \{c_1, \dots, c_n \}$ such that $\sum\limits_{i=1}^n length(c_i) * w_i$ is minimum.


\subsection{Additional Coding Methods}

Other methods such as LZW and arithmetic encoding have a more improved encoding scheme \cite{website:wiki:arithmetic-encoding, website:wiki:LZW}. Unfortunately, these $2$ schemes were patented and people should pay the owners to use them. Recently they went into public domain.

LZW is used for the GIF image format. Arithmetic encoding is currently used in WebP, an image format developed by Google \cite{website:WebP}.

LZW main idea is similar with an online hashing algorithm. Encode each character into a dictionary. Each time take the prefix string $W$ which was already encoded. Concatenate $W$ with the next character in the string and add it to dictionary of encoded strings.

Arithmetic encoding uses a probability model, encodings are values between $[0, 1]$ on the real line \cite{website:ArithmeticEncoding}.


\section{Homework}

\begin{itemize}
	\item Editor \cite{website:infoarena/editor}. Tsunami \cite{website:infoarena/tsunami}. Trompeta \cite{website:infoarena/trompeta}.
	\item Muzeu \cite{website:infoarena/muzeu}. Inv \cite{website:infoarena/inv}. Binar \cite{website:infoarena/binar}. Stergeri \cite{website:infoarena/stergeri}. 
	\item Sdo \cite{website:infoarena/sdo}. ClassicTask \cite{website:infoarena/classictask} - optional. Nrtri \cite{website:infoarena/nrtri}. Merge-K-Lists \cite{website:leetcode/merge}.
	\item SlidingWindow \cite{website:infoarena/slidingwindow}
	\item Scandura \cite{website:infoarena/scandura}
\end{itemize}

\subsection{Grading} Scoring at least $60$ points on each problem will get you maximum grade (which means a bonus of $10\%$ from your final grade). Each problem category values $0.3$ only if you place in the first $5$ people with successful submissions. An answer in class has an $0.1$ score. You can get a maximum bonus of $10/100$ points for the exam.

%subsec to cover: divide and conquer, recurrence relations, binary search trees (AVL), huffman encoding, quick sort, heapsort, hashing, graphs, mst
%some problems as backups:
%telina: have a string s(1)...s(n): choose for each i (n->1) if you keep the string or reverse it's suffix
%odd number of times in logn; elements come in pairs;
%
% ---- Bibliography ----
%
%\begin{thebibliography}{5}
%

\section{Back-up problems}

\subsection{Secv6 \cite{website:infoarena/secv6}}
\subsection{Dynamic GCD \cite{website:timus/gcd}}
\subsection{Matrice 2 \cite{website:infoarena/matrice2}}
\subsection{Cabane \cite{website:infoarena/cabane}}
\subsection{Minim2 \cite{website:infoarena/minim2}}


\clearpage
\bibliographystyle{splncs}
\bibliography{llncs}
%\end{thebibliography}

\end{document}
